Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-10-12 Manuscript/manuscript.tex
Corrected numbers of contacts in stoichiometry section and added probabilities of binding towards PO4 or carbonyls.

In addition, we estimated relative binding affinities 
of several moieties in POPC towards \ce{Ca^{2+}}. 
The dominant contribution
to the binding of \ce{Ca^{2+}} to POPC membranes comes from the phosphate group. 
Such a finding is based on the calculated
number of contacts per lipid between \ce{Ca^{2+}} and POPC oxygen atoms, 
$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O}$,
in which a contact between the two atoms is counted 
if their distance is less than or equal $0.3 \, \mathrm{nm}$ 
(robustness of this definition discussed in reference~\citenum{javanainen17}). 
The average number of contacts between \ce{Ca^{2+}} and any oxygen atom in POPC is 
$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.47 $,    % = 59.8/128 
whereas if only phosphate oxygen atoms are considered 
the average number of contacts decreases only by a tiny amount to 
$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.45 $,    % = 57.2/128 . 
and if only carbonyl oxygen atoms are considered,
this value is merely
$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.02 $.    % = (59.8-57.2)/128 . 
We corroborate this finding with the probability isodensity contours in Fig.~\ref{fig:volmaps}, 
which are easier to interpret.

@@

Internal side-note: 
However, $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$ is increased 
when carbonyl dipoles are scaled only with 0.945 
(roughly done through charge redistribution)
 to : $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} \approx 0.08 $.    % = (70.-59.8)/128 . 
This does not change the dominance of phosphates, which still interact more than 6x times more. That's good.
In addition, the states that have a contribution from carbonyl oxygens also contain a bound phosphate, 
so it can be thought of rather as a stabilizing moiety. 
In addition, the probability of states, containing sn-1:O increases from 6\% to 10\%.
The residence times change to 75ns (90\% of bound time) and max is 185ns.
That's not much, too. It certainly is not a game changer, I'd even call it to be within accuracy. 
The slight problem may be that with this tweak, there's approx. 20\%  more cations bound to the membrane,
which means, I'd need to tune the sigma parameter a bit (increase the scaling factor $f_\sigma$). 
That could detune the membrane structure \dots
Not $f_q$ -- that's set by the simulations with the surfactant. Good.
I think that such a binding detail is within the accuracy of the proposed simple correction.
It would be worth noting that the model may underestimate slightly the contribution from carbonyls due to this. 
But we can't say to what extent\dots

>>>>

In addition, we estimated relative binding affinities 
of several moieties in POPC towards \ce{Ca^{2+}}. 
The dominant contribution
to the binding of \ce{Ca^{2+}} to POPC membranes comes from the phosphate group. 
Such a finding is based on the calculated
number of contacts per lipid between \ce{Ca^{2+}} and POPC oxygen atoms, 
$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O}$,
in which a contact between the two atoms is counted 
if their distance is less than or equal $0.3 \, \mathrm{nm}$ 
(robustness of this definition discussed in reference~\citenum{javanainen17}). 
The average number of contacts per lipid between \ce{Ca^{2+}} and any oxygen atom in POPC is 
$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 $,    % = 33.8/128 
whereas if only phosphate oxygen atoms are considered 
the average number of contacts decreases only by a tiny amount to 
$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.262 $,    % = 33.5/128 . 
and if only carbonyl oxygen atoms are considered 
(phosphate oxygens contribute, but are not counted as contacts),
this value is merely
$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.066 $.    % = 8.4/128 . 
This means that calcium is in contact with a phosphate oxygen for 99\% of the time it is bound to the membrane, 
whereas only 25\% of the time accounts for contact with a carbonyl oxygen,
and purely carbonyl bound states account for less than 1\% of configurations. 
We corroborate this finding with the probability isodensity contours in Fig.~\ref{fig:volmaps}, 
which are easier to interpret.

@@

Internal side-note: 
However, $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$ is increased 
when carbonyl dipoles are scaled only with 0.945 
(roughly done through charge redistribution)
 to : $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} \approx 0.08 $.    % = (70.-59.8)/128 .

The average number of contacts per lipid between \ce{Ca^{2+}} and any oxygen atom in POPC is 
$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 \rightarrow 0.309 $,
whereas if only phosphate oxygen atoms are considered 
the average number of contacts decreases only by a tiny amount to 
$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.262 \rightarrow 0.303 $,    % = 33.5/128 . 
and if only carbonyl oxygen atoms are considered 
(phosphate oxygens contribute, but are not counted as contacts),
this value is merely
$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.066 \rightarrow 0.139 $.    % = 8.4/128 .

This means that calcium is in contact with a phosphate oxygen for 98\% of the time it is bound to the membrane, 
whereas 46\% of the time accounts for contact with a carbonyl oxygen,
and purely carbonyl bound states account for less than 2\% of configurations.

The states that have a contribution from carbonyl oxygens also contain a bound phosphate, 
so it can be thought of rather as a stabilizing moiety. 
In addition, the probability of states (among bound states), containing sn-1:O increases from 20\% to 28\% (from MSM).
Purely carbonyl-bound states make only 0.7\% and in carbonyl-dip corrected 1.4\%.
The residence times change to 75ns (90\% of bound time) and max is 185ns.
That's not much, too. It certainly is not a game changer, I'd even call it to be within accuracy. 
The slight problem may be that with this tweak, there's approx. 17\%  more cations bound to the membrane,
which means, I'd need to tune the sigma parameter a bit (increase the scaling factor $f_\sigma$). 
That could detune the membrane structure \dots
Not $f_q$ -- that's set by the simulations with the surfactant. Good.
I think that such a binding detail is within the accuracy of the proposed simple correction.
It would be worth noting that the model may underestimate slightly the contribution from carbonyls due to this. 
But we can't say to what extent\dots
